---
title: "Zf_scatac_6_TrajectoryAnalysis"
author: "zhangyl"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Perform trajectory analysis using diffusion pseudotime (DPT)

This step run in Python.

## Library package

```{python}
from __future__ import print_function
from sklearn import datasets
from time import time , sleep
from tqdm import tqdm
import numpy as np
import pandas as pd
import scanpy as sc
import scipy.sparse as sp
import scipy.spatial.distance
import scipy
from scipy.cluster.hierarchy import linkage, dendrogram , cophenet
import gc
from multiprocessing import Pool
```

## Load h5ad

```{python}
atac_spdome = sc.read_h5ad("/home/zhangyl/zbf_scATAC_v3/atac_spdome_new.h5ad")
```

## Run diffmap

```{python}
sc.pp.neighbors(atac_spdome, n_neighbors=13, n_pcs=12, use_rep='X_integrated_lsi', method='gauss')
sc.tl.diffmap(atac_spdome)
```

## Choose root and conculate pseudotime value

```{python}
Tn_mask = np.isin(atac_spdome.obs['type'], ['Epi1'])
min_stem_id = np.argmin(atac_spdome.obsm['X_diffmap'][Tn_mask, 1])
root_id = np.arange(len(Tn_mask))[Tn_mask][min_stem_id]
atac_spdome.uns['iroot'] = root_id
sc.tl.dpt(atac_spdome)
```

## Visualization 2D

```{python}
atac_spdome.obsm['X_diffmap_'] = atac_spdome.obsm['X_diffmap'][:,1:]
sc.pl.embedding(atac_spdome,'diffmap_',color=['dpt_pseudotime','type' , 'batch'])
```

## Visualization 3D and save plot

```{python}
sc.pl.embedding(atac_spdome,'diffmap_',color=['dpt_pseudotime','type', 'batch'] , dimensions=(0,1,2) ,
                projection = '3d' , save='diffmap_3D.pdf')
```

## Save the cell embedding of trajectory and pseudotime value

```{python}
np.savetxt("/home/zhangyl/zbf_scATAC_v3/diffmap_spdome_2.txt" , atac_spdome.obsm['X_diffmap_'] , 
           delimiter = '\t' , comments="")

np.savetxt("/home/zhangyl/zbf_scATAC_v3/dpt_pseudometime_spdome_2.txt" , atac_spdome.obs.dpt_pseudotime , 
           delimiter = '\t' , comments="")
```

# Fit smoothed accessibility curves for peaks of cells through pseudotime

This step run in R.

## Library package

```{r}
library(mgcv)
library(VGAM)
library(ggplot2)
library(pheatmap)
library(RColorBrewer)
```

## Find differentail accessible peaks

```{r}
markers_spdome_DEL1_new <- FindMarkers(object = atac_spdome_new,
                                       ident.1 = "DEL1",
                                       min.pct = 0.05,
                                       test.use = 'LR',
                                       latent.vars = 'uniq_nuc_frags')
markers_spdome_DEL2_new <- FindMarkers(object = atac_spdome_new,
                                       ident.1 = "DEL2",
                                       min.pct = 0.05,
                                       test.use = 'LR',
                                       latent.vars = 'uniq_nuc_frags')
markers_spdome_evl_new <- FindMarkers(object = atac_spdome_new,
                                       ident.1 = "EVL",
                                       min.pct = 0.05,
                                       test.use = 'LR',
                                      latent.vars = 'uniq_nuc_frags')
markers_spdome_ysl_new <- FindMarkers(object = atac_spdome_new,
                                      ident.1 = "E-YSL",
                                      min.pct = 0.05,
                                      test.use = 'LR',
                                      latent.vars = 'uniq_nuc_frags')

markers_spdome_evl_new$peaks <- rownames(markers_spdome_evl_new)
markers_spdome_ysl_new$peaks <- rownames(markers_spdome_ysl_new)
markers_spdome_epi2_new$peaks <- rownames(markers_spdome_epi2_new)
markers_spdome_epi1_new$peaks <- rownames(markers_spdome_epi1_new)
```

## Add pseudotime value to meta.data

```{r}
dpt_pseudotime <- read.table(file = "/home/zhangyl/zbf_scATAC_v3/dpt_pseudometime_spdome_2.txt" , quote = F , sep = "\t")
atac_spdome_new$dpt_pseudotime <- dpt_pseudotime$V1
```

## Choose cells in each trajectory

```{r}
cells = colnames(atac_spdome_new_new@assays[[assay]]@data)
cells_traj_evl <- c(rownames(atac_spdome_new@meta.data)[atac_spdome_new@meta.data$dpt_pseudotime < 0.19] , 
                    rownames(atac_spdome_new@meta.data)[atac_spdome_new@meta.data$type == "EVL"])
cells_traj_ysl <- c(rownames(atac_spdome_new@meta.data)[atac_spdome_new@meta.data$dpt_pseudotime < 0.19] , 
                    rownames(atac_spdome_new@meta.data)[atac_spdome_new@meta.data$type == "E-YSL"])
cells_traj_epi <- c(rownames(atac_spdome_new@meta.data)[atac_spdome_new@meta.data$dpt_pseudotime < 0.19] , 
                    rownames(atac_spdome_new@meta.data)[atac_spdome_new@meta.data$type == "Epi2"])
```

## Choose top 1000 differential accessibility peaks of each type to fit

```{r}
features_evl <- head(markers_spdome_evl_new$peaks , n = 1000)
features_ysl <- head(markers_spdome_ysl_new$peaks , n = 1000)
features_epi <- head(markers_spdome_epi2_new$peaks , n = 1000)

features <- data.frame("peaks" = c(features_evl , features_ysl , features_epi) , "type" = rep(c("EVL" , "YSL" , "DEL2") , each = 1000))
features <- features[duplicated(features$peaks) == 0,]
```

## Scale matrix

```{r}
atac_spdome_new <- ScaleData(atac_spdome_new , assay = "peaks" , features = features$peaks)
```

## Create a function to fit a vector generalized additive model (VGAM) for each trajectory through pseudotime

df : degrees of freedom.See VGAM.

```{r}
h <- function(object , cells = colnames(object@assays[[assay]]@data) , 
              features = rownames(object@assays[[assay]]@data) , assay = DefaultAssay(integrated_spdome) , df = 3)
{
  d <- object@assays[[assay]]@scale.data
  d <- as.data.frame(t(d[features,cells]))
  d <- cbind(d , pseudotime = object$dpt_pseudotime[cells] , stage = object$type[cells])
  ss <- seq(0,1,by = 0.01)
  dd <- data.frame(test = ss)
  for(i in features){
    fit <- vgam(data = d , d[,i] ~ sm.ns(pseudotime , df = df) , family = uninormal() , model = T)
    dd <- cbind(dd, predict(fit , data.frame(pseudotime = ss))[,1])
  }
  dd <- dd[,-1]
  dd <- as.data.frame(t(dd))
  rownames(dd) <- features
  colnames(dd) <- ss
  return(dd)
}
```

## Fit generalized additive model and create matrix

```{r}
dd_evl <- h(object = atac_spdome_new , cells = cells_traj_evl , features = features$peaks , assay = "peaks")
dd_ysl <- h(object = atac_spdome_new , cells = cells_traj_ysl , features = features$peaks , assay = "peaks")
dd_epi <- h(object = atac_spdome_new , cells = cells_traj_epi , features = features$peaks , assay = "peaks")
dd <- cbind(dd_evl , dd_ysl , dd_epi)
dd[dd > 6] = 6
```

## Draw heatmap

```{r}
anno_row <- data.frame("type" = features$type , row.names = features$peaks)
anno_colors <- list(type = c(EVL = "#e58932" , YSL = "#80598f" , DEL2 = "#c0b286"))

dp <- pheatmap(dd , color = colorRampPalette(rev(brewer.pal(n = 7, name = "Spectral")))(100) ,
               cluster_cols = F , cluster_rows = T , border_color = F , show_colnames = F,
               annotation_row = anno_row, annotation_colors = anno_colors,
               scale = "none" , show_rownames = F , cutree_rows = 4, gaps_col = c(101,202))

dp <- as.ggplot(dp)
```
