---
title: "Zf_scatac_2_IntergrateAllStage"
author: "zhangyl"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages

```{r}
library(Signac)
library(Seurat)
library(GenomicRanges)
library(future)
library(ggplot2)
library(patchwork)
library(hdf5r)
library(dplyr)
library(readr)
library(pheatmap)
library(ggrepel)
library(LSD)
library(MASS)
library(future)
library(data.table)
library(pROC)
library(SeuratDisk)
```

## Create a common set of peaks

```{r}
gr.1k <- granges(atac_1k)
gr.high <- granges(atac_high)
gr.oblong <- granges(atac_oblong)
gr.sphere <- granges(atac_sphere)
gr.dome <- granges(atac_dome)
```

## Remove peaks intersected with blacklist

The blacklist was generated from a prior dataset of ATAC-seq data obtained from zebrafish nucleosome-free genomics DNA.

```{r}
gr.1k_qc <- gr.1k[is.na(findOverlaps(gr.1k, bl.peaks, select="arbitrary"))]
gr.high_qc <- gr.high[is.na(findOverlaps(gr.high, bl.peaks, select="arbitrary"))]
gr.oblong_qc <- gr.oblong[is.na(findOverlaps(gr.oblong, bl.peaks, select="arbitrary"))]
gr.sphere_qc <- gr.sphere[is.na(findOverlaps(gr.sphere, bl.peaks, select="arbitrary"))]
gr.dome_qc <- gr.dome[is.na(findOverlaps(gr.dome, bl.peaks, select="arbitrary"))]

combine.peaks <- reduce(x = c(gr.1k_qc , gr.high_qc , gr.oblong_qc, gr.sphere_qc, gr.dome_qc))
peakwidths <-width(combine.peaks)

range(peakwidths)
```

## Create fragment objects

```{r}
frags.1k <- CreateFragmentObject(path = "~/zbf_scATAC_mtx_2/1k/aggregate_fragments.tsv.gz" , 
                                 cells = rownames(atac_1k@meta.data))
frags.high <- CreateFragmentObject(path = "~/zbf_scATAC_mtx_2/high/aggregate_fragments.tsv.gz" , 
                                   cells = rownames(atac_high@meta.data))
frags.oblong <- CreateFragmentObject(path = "~/zbf_scATAC_mtx_2/oblong/aggregate_fragments.tsv.gz" , 
                                     cells = rownames(atac_oblong@meta.data))
frags.sphere <- CreateFragmentObject(path = "~/zbf_scATAC_mtx_2/sphere/aggregate_fragments.tsv.gz" , 
                                     cells = rownames(atac_sphere@meta.data))
frags.dome <- CreateFragmentObject(path = "~/zbf_scATAC_mtx_2/dome/aggregate_fragments.tsv.gz" , 
                                   cells = rownames(atac_dome@meta.data))
```

## Create meta.data

```{r}
md.1k <- atac_1k@meta.data
md.high <- atac_high@meta.data
md.oblong <- atac_oblong@meta.data
md.sphere <- atac_sphere@meta.data
md.dome <- atac_dome@meta.data
```

## Create Seurat objects of each stage

```{r}
atac1k.counts <- FeatureMatrix(
  fragments = frags.1k , 
  features = combine.peaks , 
  cells = rownames(md.1k)
)
atac_high.counts <- FeatureMatrix(
  fragments = frags.high , 
  features = combine.peaks , 
  cells = rownames(md.high)
)
atac_oblong.counts <- FeatureMatrix(
  fragments = frags.oblong , 
  features = combine.peaks , 
  cells = rownames(md.oblong)
)
atac_sphere.counts <- FeatureMatrix(
  fragments = frags.sphere , 
  features = combine.peaks , 
  cells = rownames(md.sphere)
)
atac_dome.counts <- FeatureMatrix(
  fragments = frags.dome , 
  features = combine.peaks , 
  cells = rownames(md.dome)
)

atac_1k <- CreateChromatinAssay(atac1k.counts , fragments = frags.1k) %>% CreateSeuratObject(assay = "peaks" , meta.data = md.1k)
atac_high <- CreateChromatinAssay(atac_high.counts , fragments = frags.high) %>% CreateSeuratObject(assay = "peaks" , meta.data = md.high)
atac_oblong <- CreateChromatinAssay(atac_oblong.counts , fragments = frags.oblong) %>% CreateSeuratObject(assay = "peaks" , meta.data = md.oblong)
atac_sphere <- CreateChromatinAssay(atac_sphere.counts , fragments = frags.sphere) %>% CreateSeuratObject(assay = "peaks" , meta.data = md.sphere)
atac_dome <- CreateChromatinAssay(atac_dome.counts , fragments = frags.dome) %>% CreateSeuratObject(assay = "peaks" , meta.data = md.dome)
```

## Merge objects

```{r}
atac_merge <- merge(x = atac_1k , y = list(atac_high , atac_oblong , atac_sphere , atac_dome))
Annotation(atac_merge) <- annotation.zbf
Idents(atac_merge) <- factor(atac_merge$batch)
```

## Save object

```{r}
saveRDS(object = atac_merge , file = "~/zbf_scATAC_n/atac_merge.RDS")
```

## Convert object to h5ad

```{r}
SaveH5Seurat(atac_merge , filename = "atac_merge.h5Seurat")
Convert("atac_merge.h5Seurat" , dest = "h5ad")
```

## Integration

We integrated data based on BBKNN. This step was run in Python.

```{python}
## Library package
%matplotlib inline
%config InlineBackend.figure_format = 'retina'
import gzip
import numpy as np
import pandas as pd
import seaborn as snsNote that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

import matplotlib as mpl
import bbknn as bbknn
from collections import Counter
import matplotlib.pyplot as plt
import scanpy as sc
sc.settings.verbosity = 3             # verbosity: errors (0), warnings (1), info (2), hints (3)
sc.logging.print_header()
sc.settings.set_figure_params(dpi=80)
```

## Load data

```{python}
adata = sc.read_h5ad("/home/zhangyl/zbf_scATAC_v3/RDS/atac_merge.h5ad")
```

## Add ref/query label

```{python}
adata.obs['experimental_label'] = \
    adata.obs.batch.replace({0 : '1k_rep1', 2 : 'high', 4 : 'sphere_rep1', 8 : 'sphere-dome', 9 : 'dome_rep1',
                             1 : '1k_rep2', 5 : 'sphere_rep2', 10 : 'dome_rep2'})
adata.obs['for_ingest'] = adata.obs.experimental_label
adata.obs['refquery'] = \
    adata.obs.batch.replace({0 : 'ref', 1 : 'ref', 2 : 'ref', 4 : 'ref', 9 : 'ref',
                             5 : 'query', 8 : 'query', 10 : 'query'})
sc.pl.umap(adata, color='experimental_label' , legend_loc='on data')
sc.pl.umap(adata, color='plate' , legend_loc='on data')
```

## PCA

```{python}
sc.tl.pca(adata)
sc.pl.pca_variance_ratio(adata, log=True, n_pcs=50)
```

## Perform BBKNN

```{python}
sc.external.pp.bbknn(adata, batch_key='refquery', use_rep = "lsi2" ,n_pcs=40)
sc.tl.umap(adata)
sc.tl.leiden(adata, resolution = 0.25)
```
