---
title: "Zf_scatac_2_IntergrateAllStage"
author: "zhangyl"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages

```{r}
library(Signac)
library(Seurat)
library(GenomicRanges)
library(future)
library(ggplot2)
library(patchwork)
library(hdf5r)
library(dplyr)
library(readr)
library(pheatmap)
library(ggrepel)
library(LSD)
library(MASS)
library(future)
library(data.table)
library(pROC)
library(SeuratDisk)
```

## Create a common set of peaks

```{r}
gr.1k <- granges(atac_1k)
gr.high <- granges(atac_high)
gr.oblong_sphere <- granges(atac_oblong_sphere)
gr.sphere_dome <- granges(atac_sphere_dome)
gr.dome <- granges(atac_dome)
```

## Remove peaks intersected with blacklist

The blacklist was generated from a prior dataset of ATAC-seq data obtained from zebrafish nucleosome-free genomics DNA.

```{r}
gr.1k_qc <- gr.1k[is.na(findOverlaps(gr.1k, bl.peaks, select="arbitrary"))]
gr.high_qc <- gr.high[is.na(findOverlaps(gr.high, bl.peaks, select="arbitrary"))]
gr.oblong_sphere_qc <- gr.oblong_sphere[is.na(findOverlaps(gr.oblong_sphere, bl.peaks, select="arbitrary"))]
gr.sphere_dome_qc <- gr.sphere_dome[is.na(findOverlaps(gr.sphere_dome, bl.peaks, select="arbitrary"))]
gr.dome_qc <- gr.dome[is.na(findOverlaps(gr.dome, bl.peaks, select="arbitrary"))]

combine.peaks <- reduce(x = c(gr.1k_qc , gr.high_qc , gr.oblong_sphere_qc, gr.sphere_dome_qc, gr.dome_qc))
peakwidths <-width(combine.peaks)

range(peakwidths)
```

## Create fragment objects

```{r}
frags.1k <- CreateFragmentObject(path = "1k/aggregate_fragments.tsv.gz" , 
                                 cells = rownames(atac_1k@meta.data))
frags.high <- CreateFragmentObject(path = "high/aggregate_fragments.tsv.gz" , 
                                   cells = rownames(atac_high@meta.data))
frags.oblong_sphere <- CreateFragmentObject(path = "oblong_sphere/aggregate_fragments.tsv.gz" , 
                                     cells = rownames(atac_oblong_sphere@meta.data))
frags.sphere_dome <- CreateFragmentObject(path = "sphere_dome/aggregate_fragments.tsv.gz" , 
                                     cells = rownames(atac_sphere_dome@meta.data))
frags.dome <- CreateFragmentObject(path = "dome/aggregate_fragments.tsv.gz" , 
                                   cells = rownames(atac_dome@meta.data))
```

## Create meta.data

```{r}
md.1k <- atac_1k@meta.data
md.high <- atac_high@meta.data
md.oblong_sphere <- atac_oblong_sphere@meta.data
md.sphere_dome <- atac_sphere_dome@meta.data
md.dome <- atac_dome@meta.data
```

## Create Seurat objects of each stage

```{r}
atac1k.counts <- FeatureMatrix(
  fragments = frags.1k , 
  features = combine.peaks , 
  cells = rownames(md.1k)
)
atac_high.counts <- FeatureMatrix(
  fragments = frags.high , 
  features = combine.peaks , 
  cells = rownames(md.high)
)
atac_oblong_sphere.counts <- FeatureMatrix(
  fragments = frags.oblong_sphere , 
  features = combine.peaks , 
  cells = rownames(md.oblong_sphere)
)
atac_sphere_dome.counts <- FeatureMatrix(
  fragments = frags.sphere_dome , 
  features = combine.peaks , 
  cells = rownames(md.sphere_dome)
)
atac_dome.counts <- FeatureMatrix(
  fragments = frags.dome , 
  features = combine.peaks , 
  cells = rownames(md.dome)
)

atac_1k <- CreateChromatinAssay(atac1k.counts , fragments = frags.1k) %>% CreateSeuratObject(assay = "peaks" , meta.data = md.1k)
atac_high <- CreateChromatinAssay(atac_high.counts , fragments = frags.high) %>% CreateSeuratObject(assay = "peaks" , meta.data = md.high)
atac_oblong_sphere <- CreateChromatinAssay(atac_oblong_sphere.counts , fragments = frags.oblong_sphere) %>% CreateSeuratObject(assay = "peaks" , meta.data = md.oblong_sphere)
atac_sphere_dome <- CreateChromatinAssay(atac_sphere_dome.counts , fragments = frags.sphere_dome) %>% CreateSeuratObject(assay = "peaks" , meta.data = md.sphere_dome)
atac_dome <- CreateChromatinAssay(atac_dome.counts , fragments = frags.dome) %>% CreateSeuratObject(assay = "peaks" , meta.data = md.dome)
```

## Merge objects

```{r}
atac_merge <- merge(x = atac_1k , y = list(atac_high , atac_oblong_sphere , atac_sphere_dome , atac_dome))
Annotation(atac_merge) <- annotation.zbf
```

## Add batch

```{r}
atac_merge$batch <- atac_merge$plate
atac_merge$batch[atac_merge$plate %in% c("zf_3.0hpf_plate1", "zf_3.0hpf_plate2", "zf_3.0hpf_plate3", "zf_3.0hpf_plate4")] <- "zf_3.0hpf_batch1"
atac_merge$batch[atac_merge$plate %in% c("zf_3.0hpf_plate5", "zf_3.0hpf_plate6")] <- "zf_3.0hpf_batch2"
atac_merge$batch[atac_merge$plate %in% c("zf_3.3hpf_plate1", "zf_3.3hpf_plate2", "zf_3.3hpf_plate3", "zf_3.3hpf_plate4")] <- "zf_3.3hpf_batch1"
atac_merge$batch[atac_merge$plate %in% c("zf_3.7hpf_plate1", "zf_3.7hpf_plate2", "zf_3.7hpf_plate3", "zf_3.7hpf_plate4")] <- "zf_3.7hpf_batch1"
atac_merge$batch[atac_merge$plate %in% c("zf_3.7hpf_plate5", "zf_3.7hpf_plate6", "zf_3.7hpf_plate7", "zf_3.7hpf_plate8")] <- "zf_3.7hpf_batch2"
atac_merge$batch[atac_merge$plate %in% c("zf_4.1hpf_plate1", "zf_4.1hpf_plate2", "zf_4.1hpf_plate3", "zf_4.1hpf_plate4")] <- "zf_4.1hpf_batch1"
atac_merge$batch[atac_merge$plate %in% c("zf_4.3hpf_plate1", "zf_4.3hpf_plate2", "zf_4.3hpf_plate3", "zf_4.3hpf_plate4")] <- "zf_4.3hpf_batch1"
atac_merge$batch[atac_merge$plate %in% c("zf_4.3hpf_plate5", "zf_4.3hpf_plate6", "zf_4.3hpf_plate7", "zf_4.3hpf_plate8")] <- "zf_4.3hpf_batch2"

Idents(atac_merge) <- factor(atac_merge$batch)
```


## Normalization and linear dimensional reduction

```
atac_merge <- RunTFIDF(atac_merge)
atac_merge <- FindTopFeatures(atac_merge, min.cutoff = 'q0')
atac_merge <- RunSVD(atac_merge)

atac_merge <- RunUMAP(object = atac_merge, reduction = 'lsi', dims = 2:30)
atac_merge <- FindNeighbors(object = atac_merge, reduction = 'lsi', dims = 2:30)
atac_merge <- FindClusters(object = atac_merge, verbose = FALSE, algorithm = 3)
```

## Save object

```{r}
saveRDS(object = atac_merge , file = "atac_merge.RDS")
```

## Convert object to h5ad

```{r}
SaveH5Seurat(atac_merge , filename = "atac_merge.h5Seurat")
Convert("atac_merge.h5Seurat" , dest = "h5ad")
```

## Integration

We integrated data based on BBKNN. This step was run in Python.

```{python}
## Library package
%matplotlib inline
%config InlineBackend.figure_format = 'retina'
import gzip
import numpy as np
import pandas as pd
import seaborn as snsNote

import matplotlib as mpl
import bbknn as bbknn
from collections import Counter
import matplotlib.pyplot as plt
import scanpy as sc
sc.settings.verbosity = 3 
sc.logging.print_header()
sc.settings.set_figure_params(dpi=80)
```

## Load data

```{python}
adata = sc.read_h5ad("atac_merge.h5ad")
```

## Add ref/query label

```{python}
adata.obs['refquery'] = \
    adata.obs.batch.replace({'zf_3.0hpf_batch1' : 'ref', 'zf_3.3hpf_batch1' : 'ref', 'zf_3.7hpf_batch1' : 'ref', 'zf_4.1hpf_batch1' : 'ref', 'zf_4.3hpf_batch1' : 'ref',
                             'zf_3.0hpf_batch2' : 'query', 'zf_3.7hpf_batch2' : 'query', 'zf_4.3hpf_batch2' : 'query'})
sc.pl.umap(adata, color='batch' , legend_loc='on data')
```

## PCA

```{python}
sc.tl.pca(adata)
sc.pl.pca_variance_ratio(adata, log=True, n_pcs=50)
```

## Perform BBKNN

```{python}
sc.external.pp.bbknn(adata, batch_key='refquery', use_rep = "X_lsi" ,n_pcs=40)
sc.tl.umap(adata)
sc.tl.leiden(adata, resolution = 0.25)
sc.pl.umap(adata, color = ['batch', 'refquery', 'leiden'], wspace=0.4, cmap='Set3', save='lsi_40.pdf')
```
