---
title: "Zf_scatac_3_NMF"
author: "zhangyl"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Factorization matrix

We factorize peak-cell matrix of merged object and identify peak modules used non-nagative matrix foctorization (NMF) algorithm from Python scikit-learn toolkit.

## Library package

```{python}
from __future__ import print_function
from sklearn import datasets
from time import time , sleep
from tqdm import tqdm
import numpy as np
import pandas as pd
from sklearn.decomposition import NMF
import scanpy as sc
import scipy.sparse as sp
import scipy.spatial.distance
import scipy
from scipy.cluster.hierarchy import linkage, dendrogram , cophenet
import gc
from multiprocessing import Pool
```

## Perform NMF

```{python}
adata = sc.read_h5ad("./atac_merge.h5ad")
x = adata.X.todense()
model = NMF(n_components = 25 , init="nndsvdar" , random_state=0 , max_iter=5000)
W = model.fit_transform(x)
H = model.components_
```

## Save matrices

```{python}
np.savetxt("./W_atac_merge.txt" , W , delimiter=",")
np.savetxt("./H_atac_merge.txt" , H , delimiter=",")
```

# Create seurat redution object in R

This step was run in R.

## Library package

```{r}
library(Signac)
library(Seurat)
library(GenomicRanges)
library(future)
library(ggplot2)
library(patchwork)
library(hdf5r)
library(dplyr)
library(readr)
library(pheatmap)
library(ggrepel)
library(LSD)
library(MASS)
library(future)
library(data.table)
library(pROC)
library(SeuratDisk)
```

## Read matrix

```{r}
W_merge <- fread(file = "./W_atac_merge.txt" ,header = F)
H_merge <- fread(file = "./H_atac_merge.txt" ,header = F)

W_merge <- as.data.frame(W_merge)
rownames(W_merge) <- rownames(atac_merge@meta.data)
colnames(W_merge) <- paste("merge_module", c(1:25) , sep = '')

H_merge <- as.data.frame(t(H_merge))
rownames(H_merge) <- rownames(atac_merge@assays$peaks)
colnames(H_merge) <- paste("merge_module", c(1:25) , sep = '')
```

## Create DimReduc object

```{r}
atac_merge_nmf_reduction <- CreateDimReducObject(embeddings = as.matrix(W_merge) , 
                                                 loadings = as.matrix(H_merge) , 
                                                 assay = "peaks" , 
                                                 key = "NMF_" , 
                                                 global = T)
```

## Save object

```{r}
atac_merge <- readRDS(file = "./atac_merge.RDS")
atac_merge@reductions$NMF <- atac_merge_nmf_reduction
```

## Visualization

```{r}
DimPlot(atac_merge , label = T , group.by = "batch")
FeaturePlot(atac_merge , label = T , features = "NMF_2")
FeaturePlot(atac_merge , label = T , features = "NMF_3")
FeaturePlot(atac_merge , label = T , features = "NMF_4")
FeaturePlot(atac_merge , label = T , features = "NMF_6")
FeaturePlot(atac_merge , label = T , features = "NMF_10")

VlnPlot(atac_merge , features = c("NMF_1" , "NMF_2" , "NMF_3" , "NMF_4" , "NMF_5" , 
                                  "NMF_6" , "NMF_7" , "NMF_8" , "NMF_9" , "NMF_10" , 
                                  "NMF_11" , "NMF_12" , "NMF_13" , "NMF_14" , "NMF_15" , 
                                  "NMF_16" , "NMF_17" , "NMF_18" , "NMF_19" , "NMF_20" , 
                                  "NMF_21" , "NMF_22" , "NMF_23" , "NMF_24" , "NMF_25") , ncol = 5 ,
        group.by = "batch")

VlnPlot(atac_merge , features = c("NMF_1" , "NMF_2" , "NMF_3" , "NMF_4" , "NMF_5" , 
                                  "NMF_6" , "NMF_7" , "NMF_8" , "NMF_9" , "NMF_10" , 
                                  "NMF_11" , "NMF_12" , "NMF_13" , "NMF_14" , "NMF_15" , 
                                  "NMF_16" , "NMF_17" , "NMF_18" , "NMF_19" , "NMF_20" , 
                                  "NMF_21" , "NMF_22" , "NMF_23" , "NMF_24" , "NMF_25") , ncol = 5 ,
        group.by = "stage")
```

## Choose peaks with higher weights

```{r}
top_peaks_module2 <- head(rownames(H_merge)[order(H_merge$module2 , decreasing = T)] , n = 2000)
top_peaks_module3 <- head(rownames(H_merge)[order(H_merge$module3 , decreasing = T)] , n = 2000)
top_peaks_module4 <- head(rownames(H_merge)[order(H_merge$module4 , decreasing = T)] , n = 2000)
top_peaks_module6 <- head(rownames(H_merge)[order(H_merge$module6 , decreasing = T)] , n = 2000)
top_peaks_module10 <- head(rownames(H_merge)[order(H_merge$module10 , decreasing = T)] , n = 2000)

write.table(as.data.frame(top_peaks_module2) , file = "./increase_module_2.bed" , quote = F , sep = "\t" , row.names = F)
write.table(as.data.frame(top_peaks_module3) , file = "./increase_module_3.bed" , quote = F , sep = "\t" , row.names = F)
write.table(as.data.frame(top_peaks_module4) , file = "./increase_module_4.bed" , quote = F , sep = "\t" , row.names = F)
write.table(as.data.frame(top_peaks_module6) , file = "./increase_module_6.bed" , quote = F , sep = "\t" , row.names = F)
write.table(as.data.frame(top_peaks_module10) , file = "./decrease_module_10.bed" , quote = F , sep = "\t" , row.names = F)
```

## Choose random peaks from background

```{r}
set.seed(2022)
random_peaks <- sample(x = rownames(atac_merge@assays$peaks@data) , size = 2000 , replace = F)
write.table(as.data.frame(random_peaks) , file = "./random_2000_peaks.bed" , quote = F , sep = "\t" , row.names = F)
```

## Annotation

```{r}
#### anno_module2
peakanno_module2 <- annoPeaks(peaks = StringToGRanges(top_peaks_module2) , annoData = annotation.zbf , select = "bestOne" , 
                              bindingType = c("fullRange"))

top_peaks_module2_peak <- as.data.frame(cbind(top_peaks_module2 , paste("X" , c(1:length(top_peaks_module2)) , sep = "")))
colnames(top_peaks_module2_peak) <- c("peaks" , "peak")
ad <- as.data.frame(peakanno_module2@elementMetadata)[,c(1,11,13,14)]
top_peaks_module2_anno <- merge(x = top_peaks_module2_peak , y = ad , by = "peak" , all.x = T , all.y = F , sort = T)
top_peaks_module2_anno <- top_peaks_module2_anno[,c(7,2:6,8:10)]
top_peaks_module2_anno$rank <- as.numeric(substr(top_peaks_module2_anno$peak , start = 2 , stop = 6))

top_peaks_module2_anno <- (top_peaks_module2_anno[order(top_peaks_module2_anno$rank),])
peakanno_module2 <- StringToGRanges(top_peaks_module2_anno$peaks)
peakanno_module2@elementMetadata <- DataFrame(top_peaks_module2_anno[,c(3:5)])

write.table(peakanno_module2 , file = "./todo2_decrease_module_2_annotation.bed" , quote = F , sep = "\t" , row.names = F)
```
